---
title: "RplotterPkg::create_density_plot() demos"
author: "Rick Dean"
date: "12/15/2019"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    css: ../style.css
---

```{r, message=FALSE, warning=FALSE}
library(knitr)
library(data.table)
library(rlang)
library(ggplot2)
library(dplyr)
library(RplotterPkg)
```

```{r,setup, include=FALSE, eval=TRUE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE, message = FALSE,  warning = FALSE, fig.width = 12, fig.height = 9)
```

# Demos

## 1 Introduction
The following R scripts provides some density plot examples of [RplotterPkg::create_density_plot()](https://github.com/deandevl/RplotterPkg).


## 2 Simple density function
1. Air quality data:
```{r}
str(airquality)
```

2. Plot the air quality data's estimated density:
```{r, fig.height=9, fig.width=12}
RplotterPkg::create_density_plot(
  df = airquality,
  aes_x = "Ozone",
  plot_obs = TRUE,
  rot_y_tic_label = TRUE,
  x_limits = c(-40,200),
  x_major_breaks = seq(-40,200,20)
)
```

## 3 Kernel Density Estimation(KDE)
Wikipedia provides a [KDE definition](https://en.wikipedia.org/wiki/Kernel_density_estimation). The following sample of data and plots are  provided to illustrate the KDE:

1. The data:
```{r}
data_df <-data.frame(
  x = c(-2.1, -1.3, -0.4, 1.9, 5.1, 6.2)
)
```
2. A histogram of the data:
```{r, fig.height=9, fig.width=12}
RplotterPkg::create_histogram_plot(
  df = data_df,
  bar_fill = "blue",
  bar_alpha = 0.2,
  bar_size = 1.2,
  aes_x = "x",
  bin_breaks = c(-4,-2,0,2,4,6,8),
  x_limits = c(-4,10),
  x_major_breaks = seq(-4,10,2),
  y_limits = c(0.0, 2.5),
  y_major_breaks = seq(0.0, 2.5, 0.5),
  title = "Histogram of Wikipedia Kernel Density Estimation Data",
  x_title = "X",
  bar_labels = TRUE,
  plot_obs = TRUE,
  plot_obs_color = "darkorange"
)
```
3. A density plot of the data:
```{r, fig.height=9, fig.width=12}
RplotterPkg::create_density_plot(
  df = data_df,
  aes_x = "x",
  bw = 2,
  x_limits = c(-10,12),
  x_major_breaks = seq(-10,12,2),
  title = "Kernel Density Estimation of Wikipedia Data",
  x_title = "X",
  density_fill = "black",
  density_alpha = 0.2,
  plot_obs = TRUE,
  plot_obs_color = "darkorange"
)
```

## 4 Setting the aes_x and aes_fill aesthetics.
Show the density distribution for a continuous variable across categorical levels of another variable by assigning aes_x and aes_fill.

<div class="task">
  Task: The dataset `ggplot2::midwest` has a continuous variable called *area* and a categorical variable called *state*.  Show an overlapped density distribution of *area* for 437 counties across 5 states from the midwest dataset by setting <code class="cd">aes_x = "area"</code>  and <code class="cd">aes_fill = "state"</code>.
</div> 

```{r}
str(ggplot2::midwest)
```


```{r, fig.height=9, fig.width=12}
midwest_df <- ggplot2::midwest %>% 
  mutate(area = area * 1000)

RplotterPkg::create_density_plot(
  df = midwest_df,
  aes_x = "area",
  aes_fill = "state",
  title = "Distribution of area",
  subtitle = "437 counties from midwest dataset",
  x_title = "Area",
  rot_y_tic_label = TRUE,
  density_color = "black",
  density_alpha = 0.5,
  palette_colors = c("red","green","blue","gold","black")
)
```

## 5 Random Standard Normal Distribution with area shading
1. Create a data frame with a 1000 random observations from a standard normal distributions:
```{r}
set.seed(42)
data_df <- data.frame(
  x = stats::rnorm(n = 1000)
)
```
2. Create a density plot of the random data with area shading at cumulative probabilities of 2.5% and 97.5%:
```{r}
normal_plot <- RplotterPkg::create_density_plot(
  df = data_df,
  aes_x = "x",
  title = "Random Normal Distribution",
  subtitle = "probability area shading",
  x_limits = c(-4,4),
  x_major_breaks = seq(-4,4,1),
  y_limits = c(0.0, 0.5),
  y_major_breaks = seq(0.0, 0.5, 0.1),
  rot_y_tic_label = TRUE,
  cum_prob = c(0.025, 0.975),
  area_colors = c("gray", "green", "gray"),
  #cum_prob = 0.975,
  #cum_colors = c("green", "gray"),
  area_quantile_line_color = "blue"
)
normal_plot
```
