---
title: "RplotterPkg::create_scatter_plot() demos"
author: "Rick Dean"
date: "12/18/2019"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    css: ../style.css
---

```{r, message=FALSE, warning=FALSE}
library(here)
library(ggplot2)
library(rlang)
library(ggrepel)
library(grid)
library(gtable)
library(dplyr)
library(readr)
library(modelr)
library(RplotterPkg)

here::here()
```

```{r,setup, include=FALSE, eval=TRUE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE, message=FALSE, warning = FALSE, fig.width = 10, fig.height = 6)
```

# Demos

## 1 Introduction
The following R scripts provides some scatter plot examples of [RplotterPkg::create_scatter_plot()](https://github.com/deandevl/RplotterPkg).

## 2 Setting the `aes_x` and `aes_y` aesthetics.

The *modelr::sim1* dataset has both a continuous independent variable `x` and continuous dependent variable `y`.

```{r}
str(modelr::sim1)
```
<div class="task">
  Task: Set `aes_x` = "x" and `aes_y` = "y" from the *modelr::sim1* dataset and plot the points with axis' defaults.
</div>

```{r}
RplotterPkg::create_scatter_plot(
  df = modelr::sim1,
  aes_x = "x",
  aes_y = "y",
  title = "X vs Y",
  center_titles = TRUE,
  subtitle = "dataset: sim1",
  x_title = "X",
  y_title = "Y",
  rot_y_tic_label = TRUE,
  pts_color = "green",
  pts_fill = "blue",
  pts_size = 5.0,
  pts_stroke = 2.0,
  pts_line_alpha = 0.8
)
```

<div class="task">
  Task: Define a linear model for the *modelr::sim1* dataset.
</div>

```{r}
model_indep_sim1 <- lm(y ~ x, data = modelr::sim1)
summary(model_indep_sim1)
```
<div class="task">
  Task: Compute the predictions from the linear model.
</div>

```{r}
sim1_pred_df <- modelr::add_predictions(data = modelr::sim1, model = model_indep_sim1)
str(sim1_pred_df)
```

<div class="task">
  Task: Plot the x/y points and their predictions from the model by setting `df` = `sim1_pred_df`, aesthetic variables `aes_x` = "x", `aes_y` = "y". Add a *geom_line* that plots the predicted line with  <code class="cd">aes(y = pred)</code>.
</div>

```{r}
RplotterPkg::create_scatter_plot(
  df = sim1_pred_df,
  aes_x = "x",
  aes_y = "y",
  title = "Linear model for sim1 data",
  subtitle = "model: lm(y ~ x) in red",
  center_titles = TRUE,
  rot_y_tic_label = TRUE,
  x_limits = c(1,10),
  x_major_breaks = seq(from = 0, to = 10, by = 1),
  y_limits = c(0,30),
  y_major_breaks = seq(from = 0, to = 30, by = 2),
  show_minor_grids = FALSE,
  pts_color = "green",
  pts_fill = "blue",
  pts_size = 5.0,
  pts_stroke = 2.0,
  pts_line_alpha = 0.8
) + geom_line(aes(y = pred), linetype = "dashed", color = "red", size = 1.5)
```

<div class="task">
  Task: Compute residuals of the linear model <code class="cd">lm(y ~ x)</code>  for the *modelr::sim1* dataset.
</div>

```{r}
model_indep_sim1 <- lm(y ~ x, data = modelr::sim1)
residuals_sim1_df <- sim1 %>%
  modelr::add_residuals(model_indep_sim1, var = "Residual")
str(residuals_sim1_df)
```

<div class="task">
  Task: Plot the residuals by setting `df` = `residuals_sim1_df`, `aes_x` = "x", and `aes_y` = "Residual".
</div>

```{r}
RplotterPkg::create_scatter_plot(
  df = residuals_sim1_df,
  aes_x = "x",
  aes_y = "Residual",
  title = "Residuals for sim1 dataset",
  subtitle = "model: lm(y ~ x)",
  center_titles = TRUE,
  rot_y_tic_label = TRUE,
  x_limits = c(1,10),
  x_major_breaks = seq(from = 1, to = 10, by = 1),
  y_limits = c(-5,5),
  y_major_breaks = seq(from = -5, to = 5, by = 1),
  show_minor_grids = FALSE,
  pts_fill = "blue",
  pts_size = 2,
  pts_line_alpha = 0.6,
  bold_y = 0.0
)
```

## 3 Setting the `aes_x`, `aes_y`, and `aes_fill` aesthetics.

The *modelr::sim3* dataset has a continuous independent variable `x1`, a categorical independent variable `x2` and a continuous dependent variable `y`.

```{r}
str(modelr::sim3)
```
`x1` and `x2` are independent variables where `x2` is a factor variable with 4 levels.

<div class="task">
  Task: Define a linear model for the *modelr::sim3* dataset involving `x1` and `x2` as independent variables.
</div>

```{r}
model_indep_sim3 <- lm(y ~ x1 + x2, data = modelr::sim3)
summary(model_indep_sim3)
```
<div class="task">
  Task: Compute residuals of the linear model `model_indep_sim3`. 
</div>

```{r}
residuals_sim3_df <- modelr::sim3 %>%
  modelr::add_residuals(model_indep_sim3, var = "Residual")
str(residuals_sim3_df)
```

<div class="task">
  Task: Plot `x1` vs residuals across levels of `x2` by setting `df` = `residuals_sim3_df`, `aes_x` = "x1", `aes_y` = "Residual" and `aes_fill` = "x2".
</div>

```{r}
RplotterPkg::create_scatter_plot(
  df = residuals_sim3_df,
  aes_x = "x1",
  aes_y = "Residual",
  aes_fill = "x2",
  title = "Residuals across x2 levels for sim3 dataset",
  subtitle = "model: lm(y ~ x1 + x2)",
  center_titles = TRUE,
  rot_y_tic_label = TRUE,
  x_limits = c(1,10),
  x_major_breaks = seq(from = 1, to = 10, by = 1),
  y_limits = c(-5,5),
  y_major_breaks = seq(from = -5, to = 5, by = 1),
  show_minor_grids = FALSE,
  pts_size = 3,
  bold_y = 0.0
)
```

## 4 Setting the `aes_x` and `aes_y` aesthetics across factor levels.

<div class="task">
  Task: Repeat the above residuals plot from the *modelr::sim3* dataset but put `x1` vs `Residual` in separate plots/panels across levels of `x2`. Remove the `aes_fill` assignment and set `factor_var` to "x2". Put each plot of `x1` vs `Residual` in separate panels/levels of `x2` on one row (set columns = 4) using <code class="cd">RplotterPkg::multi_scatter_plot()</code>.
</div>

```{r, fig.height = 8, fig.width = 12}
plot_sim3_panels_resid <- RplotterPkg::multi_scatter_plot(
  df = residuals_sim3_df,
  factor_var = "x2",
  factor_x = "x1",
  columns = 4,
  aes_y = "Residual",
  title = "Residuals across x2 levels for modelr::sim3 dataset",
  subtitle = "model: lm(y ~ x1 + x2)",
  x_limits = c(1,10),
  x_major_breaks = seq(from = 1, to = 10, by = 1),
  y_limits = c(-5,5),
  y_major_breaks = seq(from = -5, to = 5, by = 1),
  show_minor_grids = FALSE,
  bold_y = 0.0,
  col_width = 3
)
```

<div class="task">
  Task: Define a linear independent and interactive model for the *modelr::sim3* dataset.
</div>

```{r}
model_indep_sim3 <- lm(y ~ x1 + x2, data = modelr::sim3)
model_inter_sim3 <- lm(y ~ x1 * x2, data = modelr::sim3)
```
<div class="task">
  Task: Compute the residuals of both models from *modelr::sim3* by using <code class="cd">modelr::gather_residuals()</code> 
</div>

```{r}
residuals_models_sim3_df <- modelr::sim3 %>%
  modelr::gather_residuals(model_indep_sim3, model_inter_sim3, .resid = "Residual", .model = "Model")
str(residuals_models_sim3_df)
```

<div class="task">
  Task: Plot the residuals from each model by setting `df` = `residuals_models_sim3_df`, `factor_x` = "y", `aes_y` = "Residual", `factor_var` = "Model" for a separate panel display for each model level.
</div>

```{r, fig.height=11, fig.width=8}
plot_sim3_indep_inter <- multi_scatter_plot(
  df = residuals_models_sim3_df,
  factor_var = "Model",
  factor_x = "y",
  columns = 1,
  aes_y = "Residual",
  title = "Residuals of linear models for modelr::sim3 dataset",
  subtitle = "models: lm(y ~ x1 + x2), lm(y ~ x1 * x2)",
  x_limits = c(1,10),
  x_major_breaks = seq(from = 1, to = 10, by = 1),
  y_limits = c(-5,5),
  y_major_breaks = seq(from = -5, to = 5, by = 1),
  show_minor_grids = FALSE,
  bold_y = 0.0,
  col_width = 8.5
)
```


<div class="task">
  Task: From the *ggplot2::mpg* plot `cty` vs `hwy` across levels of car compacts using <code class="cd">RplotterPkg::multi_scatter_plot()</code> with the plots under a single column (i.e. columns = 1). Label the points with the model name.
</div>  

```{r,fig.width=12, fig.height=11}
compact_df <- ggplot2::mpg %>% 
  dplyr::filter(class == "compact" | class == "subcompact")
plot_mtcars <- RplotterPkg::multi_scatter_plot(
  df = compact_df,
  factor_var = "class",
  factor_x = "cty",
  columns = 1,
  aes_y = "hwy",
  aes_label = "model",
  x_limits = c(10,40),
  x_major_breaks = seq(10,40,5),
  y_limits = c(15,50),
  y_major_breaks = seq(15,50,5),
  title = "City vs Highway MPG across Compacts",
  subtitle = "source - ggplot2::mpg",
  x_title = "City MPG",
  y_titles = c("Highway MPG", "Highway MPG"),
  rot_y_tic_label = TRUE,
  pts_size = 2.5,
  pts_fill = "gold",
  show_minor_grids = FALSE,
  row_height = 5,
  col_width = 12
)
```

<div class="task">Task: Plot the *gapminder* dataset of `lifeExp` vs `pop` across `continent` and `year` using <code class="cd">RplotterPkg::multi_scatter_plot()</code>.</div>

```{r}
gapminder_df <- gapminder::gapminder %>% 
  filter(year > 1987 & continent %in% c("Africa", "Americas", "Asia", "Europe")) %>%  
  droplevels() %>% 
  mutate(year = factor(year)) %>% 
  mutate(pop = pop/1e+6) %>% 
  rename(
    Year = year,
    Population = pop,
    Life_Exp = lifeExp,
    Continent = continent
  )
str(gapminder_df)
```

```{r, fig.width=13, fig.height=10}
gapminder_plot <- RplotterPkg::multi_scatter_plot(
  df = gapminder_df,
  factor_var = "Year",
  factor_x = "Life_Exp",
  aes_y = "Population",
  aes_fill = "Continent",
  x_limits = c(20,90),
  x_major_breaks = seq(from = 20, to = 90, by = 5),
  y_limits = c(0,300),
  y_major_breaks = seq(from = 0, to = 300, by = 50),
  title = "Life expectancy vs population(millions) across continents and years",
  subtitle = "dataset: gapminder",
  pts_size = 3,
  pts_line_alpha = 0.6,
  col_width = 5
)
```

<div class="task">
  Task: From the *ggplot2::economics* dataset plot a time series of unemployment from 2000 to 2014. Accept the defaults for `x_date_labels`, and `x_major_date_breaks` in <code class="cd">RplotterPkg::create_scatter_plot()</code> for a time based x axis.
</div>  

```{r}
str(ggplot2::economics)
```

```{r, fig.height=7,fig.width=14}
RplotterPkg::create_scatter_plot(
  df = ggplot2::economics,
  aes_x = "date",
  aes_y = "unemploy",
  x_title = "Date",
  y_title = "Unemployment",
  rot_y_tic_label = TRUE,
  title = "US Monthly Unemployment",
  subtitle = "2000 to 2014 (in thousands)",
  pts_color = "blue",
  line_color = "violet",
  connect = TRUE
)
```

<div class="task">
  Task: Repeat the above plot with values for `x_date_labels`, and `x_major_date_breaks` in a Year-month format. Do not draw the point symbols.
</div>

```{r, fig.width=14}
RplotterPkg::create_scatter_plot(
  df = ggplot2::economics,
  aes_x = "date",
  aes_y = "unemploy",
  x_title = "Date",
  y_title = "Unemployment",
  rot_y_tic_label = TRUE,
  title = "US Monthly Unemployment",
  subtitle = "2000 to 2014 (in thousands)",
  line_color = "violet",
  show_pts = FALSE,
  connect = TRUE,
  x_date_labels = "%Y-%b",
  x_major_date_breaks = "5 year"
) + geom_hline(aes(yintercept = 8000), color = "red", linetype = "dashed", lwd = 1)
```

<div class="task">
  Task: Repeat the above plot and set the x axis breaks as a sequence of years.
</div>  

```{r, fig.width=14}
RplotterPkg::create_scatter_plot(
  df = ggplot2::economics,
  aes_x = "date",
  aes_y = "unemploy",
  x_title = "Date",
  y_title = "Unemployment",
  rot_y_tic_label = TRUE,
  title = "US Monthly Unemployment",
  subtitle = "1967 to 2014 (in thousands)",
  line_color = "violet",
  connect = TRUE,
  x_date_labels = "%Y",
  x_major_breaks = seq.Date(from = as.Date("1964-01-01"), to = as.Date("2020-01-01"), by = "4 year"),
  show_minor_grids = FALSE
) + geom_hline(aes(yintercept = 8000), color = "red", linetype = "dashed", lwd = 1)
```


<div class="task">
  Task: Read the *yosemite_temps.csv* file.
</div>  

```{r}
yosemite_df <- readr::read_csv("../data/yosemite_temp.csv")
str(yosemite_df)

```
Convert `yosemite_df` time as POSIXct.
```{r}
yosemite_df <- yosemite_df %>% 
  mutate(time_alt = as.POSIXct(time, format = "%Y-%m-%d %H:%M"))
str(yosemite_df)
```

<div class="task">
  Task: Plot the datetime data of Yosemite temperatures recorded every 5 minutes.
</div>  

```{r}
RplotterPkg::create_scatter_plot(
  df = yosemite_df,
  title = "Yosemite Tempertures",
  subtitle = "Every 5 minutes on 2017-05-01 for 10 hours",
  x_title = "Time Hr:Min",
  y_title = "Temperature",
  rot_y_tic_label = TRUE,
  aes_x = "time_alt",
  aes_y = "y",
  y_limits = c(0,30),
  y_major_breaks = seq(from = 0, to = 30, by = 5),
  x_date_labels = "%H:%M",
  x_major_date_breaks = "30 min",
  connect = TRUE
)
```

