---
title: "RplotterPkg::create_box_plot() demos"
author: "Rick Dean"
date: "12/15/2019"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    css: ../style.css
---

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(data.table)
library(dplyr)
library(rlang)
library(ggrepel)
library(socviz)
library(RplotterPkg)
```

```{r,setup, include=FALSE, eval=TRUE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 8)
```

# Demos

## 1 Introduction

The following R scripts provides some box plot examples of [RplotterPkg::create_box_plot()](https://github.com/deandevl/RplotterPkg).
The dataset being used is from socviz::organdata.

```{r}
data("organdata")
```

## 2 Setting the aes_x and aes_y aesthetic arguments.
Showing outliers labeled with row numbers.

```{r, fig.width=10, fig.height=10}
RplotterPkg::create_box_plot(
  df = organdata,
  aes_x = "country",
  aes_y = "donors",
  y_limits = c(0,40),
  y_major_breaks = seq(from = 0, to = 40, by = 5),
  title = "Organ Donation Rate per Million",
  subtitle = "Years 1991 to 2002; outliers labeled with row numbers",
  center_titles = TRUE,
  x_title = "Country",
  y_title = "Donor Rate",
  box_color = "purple",
  box_size = 0.8,
  coord_flip = TRUE,
  rot_y_tic_label = TRUE,
  show_outliers = TRUE,
  ol_color = "red",
  ol_size = 1.5,
  order_by_median = TRUE
)
```

## 3 Setting the aes_x and aes_y along with the aes_fill aesthetics arguments.
Not showing outliers which is the default.
```{r, fig.width=10, fig.height=10}
df <- organdata %>% 
  filter(!is.na(world))

RplotterPkg::create_box_plot(
  df = df,
  aes_x = "country",
  aes_y = "donors",
  aes_fill = "world",
  y_limits = c(0,40),
  y_major_breaks = seq(from = 0, to = 40, by = 5),
  title = "Organ Donation Rate per Million",
  subtitle = "Years 1991 to 2002; no outliers",
  center_titles = TRUE,
  x_title = "Country",
  y_title = "Donor Rate",
  box_color = "purple",
  box_size = 0.8,
  coord_flip = TRUE,
  rot_y_tic_label = TRUE,
  order_by_median = TRUE
)
```

## 4 Labelling outliers with donor rate values

```{r, fig.width=10, fig.height=10}
RplotterPkg::create_box_plot(
  df = organdata,
  aes_x = "country",
  aes_y = "donors",
  y_limits = c(0,40),
  y_major_breaks = seq(from = 0, to = 40, by = 5),
  title = "Organ Donation Rate per Million",
  subtitle = "Outliers labelled by donor rate",
  center_titles = TRUE,
  x_title = "Country",
  y_title = "Donor Rate",
  box_color = "purple",
  box_size = 0.8,
  box_fill = "gold",
  coord_flip = TRUE,
  rot_y_tic_label = TRUE,
  label_outlier_var = "donors",
  label_outlier_color = "blue",
  order_by_median = TRUE
)
```

